### 2021-08-25 README.Rmd
Correcting the df name in Readme.

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# yowie <a href='https://numbats.github.io/yowie/'><img src='man/figures/logo.png' align="right" height="138.5" /></a>

<!-- badges: start -->
[![R-CMD-check](https://github.com/numbats/yowie/workflows/R-CMD-check/badge.svg)](https://github.com/numbats/yowie/actions)
<!-- badges: end -->

The goal of `yowie` is to provide longitudinal wages data sets along with several demographic variables of Americans from [the National Longitudinal Survey of Youth (NLSY79)](https://www.nlsinfo.org/content/cohorts/nlsy79) from Round 1 (the survey year 1979) to Round 28 (the survey year 2018). The NLSY79 is a longitudinal project held by the U.S. Bureau of Labor Statistics that follows the lives of a sample of American youth born between 1957-64. The cohort originally included 12,686 respondents ages 14-22 when first interviewed in 1979. There are three data sets provided in this package:

-  `wages_hs`: The wages data from the cohort whose highest grade completed is up to 12th grade.
-  `wages_hs_do`: The wages data of the high school dropouts.
-  `demog_nlsy79`: The demographic data of the NLSY79 cohort.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("numbats/yowie")
```

## Example

Here is the example of plotting some samples of subjects in `wages_hs` using `brolgar` (Tierney, Cook, and Prvan, 2020).

```{r, message=FALSE}
library(yowie)
library(brolgar)
library(ggplot2)

set.seed(20210217)

ggplot(wages_hs, 
       aes(x = year,
                y = mean_hourly_wage,
                group = id)) +
  geom_line(alpha = 0.7) +
  facet_sample() +
  ylab("mean hourly wage") +
  theme_bw()
```

## About the package name

A [yowie](https://en.wikipedia.org/wiki/Yowie) is a mythical creature in Australian foklore akin to Big Foot or Sasquatch in north America. In terms of the package it can be considered an acronym: Years Of Wages to Investigate and Explore.