### 2021-09-28 data-raw/data-preprocessing.Rmd
change pkgdown from github actions to local, added the processing of raw to input data

The data cleaning step is conducted using the suite of packages in `tidyverse`.

```{r load-pkgs}
library(tidyverse)
select <- dplyr::select
```

# Reading the data

```{r raw-data, eval=FALSE, include=FALSE}
source(here::here("data-raw/NLSY79/NLSY79.R"))
```

<details class="source">
<summary>Click here for the **source file to read data** (note this is quite long)</summary>
The code below is provided by the NLSY79 database to do the
reading and initial cleaning of data.

```{r raw-data-rmd, code=readLines(here::here("data-raw/NLSY79/NLSY79.R"))}
```

</details>
The above source code creates a data set `new_data_qnames`.
As shown below, the column names contain information on the job number
(HRP1 = job 1, HRP2 = job 2, ..., HRP5 = job 5) and the year information.

```{r untidy-data, echo=-1}
options(width=70)
new_data_qnames %>%
  select(CASEID_1979,
         starts_with("HRP") &
           ends_with(c("1979", "1980")),
         everything()) %>%
  str(list.len = 10)
```

# Demographic data pre-processing
## Tidy the date of birth data

```{r dob-tidy}
dob_tidy <- new_data_qnames %>%
  select(CASEID_1979,
                starts_with("Q1-3_A~")) %>%
  mutate(dob_year = case_when(
                      # if the years recorded in both sets match, take 79 data
                      `Q1-3_A~Y_1979` == `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`,
                      # if the year in the 81 set is missing, take the 79 data
                      is.na(`Q1-3_A~Y_1981`) ~ `Q1-3_A~Y_1979`,
                      # if the sets don't match for dob year, take the 79 data
                      `Q1-3_A~Y_1979` != `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`),
         dob_month = case_when(
                      # if months recorded in both sets match, take 79 data
                      `Q1-3_A~M_1979` == `Q1-3_A~M_1981` ~ `Q1-3_A~M_1979`,
                      # if month in 81 set is missing, take the 79 data
                      is.na(`Q1-3_A~M_1981`) ~ `Q1-3_A~M_1979`,
                      # if sets don't match for dob month, take the 79 data
                      `Q1-3_A~M_1979` != `Q1-3_A~M_1981` ~ `Q1-3_A~M_1979`),
         # flag if there is a conflict between dob recorded in 79 and 81
         dob_conflict = case_when(
                      (`Q1-3_A~M_1979` != `Q1-3_A~M_1981`) & !is.na(`Q1-3_A~M_1981`)
                      ~ TRUE,
                      (`Q1-3_A~Y_1979` != `Q1-3_A~Y_1981`) & !is.na(`Q1-3_A~Y_1981`)
                      ~ TRUE,
                      (`Q1-3_A~Y_1979` == `Q1-3_A~Y_1981`) &
                        (`Q1-3_A~M_1979` == `Q1-3_A~M_1981`) ~ FALSE,
                      is.na(`Q1-3_A~M_1981`) | is.na(`Q1-3_A~Y_1981`) ~ FALSE)) %>%
  select(CASEID_1979,
         dob_month,
         dob_year,
         dob_conflict)
```

## Tidy the sex and race data

```{r demog-tidy}
demog_tidy <- categories_qnames %>%
  select(CASEID_1979,
         SAMPLE_RACE_78SCRN,
         SAMPLE_SEX_1979) %>%
  rename(gender = SAMPLE_SEX_1979,
         race = SAMPLE_RACE_78SCRN)
```

## Tidy the grade completed in each year

```{r demog-ed}
demog_education <- new_data_qnames %>%
  as_tibble() %>%
  # in 2018, the variable's name is Q3-4_2018, instead of HGC_2018
  rename(HGC_2018 = `Q3-4_2018`) %>%
  select(CASEID_1979,
                starts_with("HGCREV"),
                "HGC_2012",
                "HGC_2014",
                "HGC_2016",
                "HGC_2018") %>%
  pivot_longer(!CASEID_1979,
               names_to = "var",
               values_to = "grade") %>%
  separate("var", c("var", "year"), sep = -4) %>%
  filter(!is.na(grade)) %>%
  select(-var)
```

### Get the highest year completed

```{r tidy-hgc}
highest_year <- demog_education %>%
  group_by(CASEID_1979) %>%
  mutate(hgc_i = max(grade)) %>%
  filter(hgc_i == grade) %>%
  filter(year == first(year)) %>%
  rename(yr_hgc = year) %>%
  select(CASEID_1979, yr_hgc, hgc_i) %>%
  ungroup() %>%
  mutate(hgc = case_when(hgc_i == 0 ~ "NONE",
                         hgc_i == 1 ~ "1ST GRADE",
                         hgc_i == 2 ~ "2ND GRADE",
                         hgc_i == 3 ~ "3RD GRADE",
                         hgc_i >= 4 & hgc_i <= 12 ~ paste0(hgc_i,"TH GRADE"),
                         hgc_i == 13 ~ "1ST YEAR COL",
                         hgc_i == 14 ~ "2ND YEAR COL",
                         hgc_i == 15 ~ "3RD YEAR COL",
                         hgc_i == 95 ~ "UNGRADED",
                         TRUE ~ paste0((hgc_i - 12), "TH YEAR COL")))

```
```{r full-demog}
full_demographics <- full_join(dob_tidy, demog_tidy, by = "CASEID_1979") %>%
  full_join(highest_year, by = "CASEID_1979") %>%
  rename("id" = "CASEID_1979") %>%
  mutate(age_1979 = 1979 - (dob_year + 1900))

```
```{r tidy-hours}
# make a list for years where we used the "QES-52A"
year_A <- c(1979:1987, 1993)
# function to get the hour of work
get_hour <- function(year) {
  if(year %in% year_A){
    temp <- new_data_qnames %>%
      select(CASEID_1979,
                    starts_with("QES-52A") &
                      ends_with(as.character(year)))}
  else{
    temp <- new_data_qnames %>%
      select(CASEID_1979,
                    starts_with("QES-52D") &
                      ends_with(as.character(year)))}
  temp %>%
    pivot_longer(!CASEID_1979,
                 names_to = "job",
                 values_to = "hours_work") %>%
    separate("job", c("job", "year"), sep = -4) %>%
    mutate(job = paste0("job_", substr(job, 9, 10))) %>%
    rename(id = CASEID_1979)
}

# getting the hours of work of all observations
years <- c(1979:1994, seq(1996, 2018, by = 2))
hours_all <- map_dfr(years, get_hour)

```
```{r tidy-rate}
get_rate <- function(year) {
  new_data_qnames %>%
    select(CASEID_1979,
                  starts_with("HRP") &
                    ends_with(as.character(year))) %>%
    pivot_longer(!CASEID_1979, names_to = "job", values_to = "rate_per_hour") %>%
    separate("job", c("job", "year"), sep = -4) %>%
    mutate(job = paste0("job_0", substr(job, 4, 4))) %>%
    rename(id = CASEID_1979)
}

rates_all <- map_dfr(years, get_rate)

# join hours and rates variable
hours_wages <- left_join(rates_all,
                         hours_all,
                         by = c("id", "year", "job")) %>%
  # set the 0 value in rate_per_hour as NA
  mutate(rate_per_hour = ifelse(rate_per_hour == 0, NA, rate_per_hour),
         hours_work = ifelse(hours_work == 0 | hours_work > 84, NA, hours_work))

```
```{r tidy-nojob}
# calculate number of jobs that a person has in one year
no_job <- hours_wages %>%
  filter(!is.na(rate_per_hour)) %>%
  group_by(id, year) %>%
  summarise(no_jobs = length(rate_per_hour))

# filter the observations with available rate per hour
eligible_wages <- hours_wages %>%
  filter(!is.na(rate_per_hour)) %>%
  left_join(no_job, by = c("id", "year"))

# calculate the mean_hourly_wage
# flag1 = code 1 for weighted mean
# code 0 for arithmetic mean
mean_hourly_wage <- eligible_wages %>%
  group_by(id, year) %>%
  #calculate the weighted mean if the number of jobs > 1
  mutate(wages = ifelse(no_jobs == 1, rate_per_hour/100,
                        weighted.mean(rate_per_hour, hours_work, na.rm = TRUE)/100)) %>%
  #give the flag if it the weighted mean
  mutate(flag1 = ifelse(!is.na(wages) & no_jobs != 1, 1,
                        0)) %>%
  #calculate the arithmetic mean for the na
  mutate(wages = ifelse(is.na(wages), mean(rate_per_hour)/100,
                        wages)) %>%
  group_by(id, year) %>%
  summarise(wages = mean(wages),
            total_hours = sum(hours_work),
            number_of_jobs = mean(no_jobs),
            flag1 = mean(flag1)) %>%
  mutate(year = as.numeric(year)) %>%
  ungroup() %>%
  rename(mean_hourly_wage = wages) %>%
  mutate(is_wm = ifelse(flag1 == 1, TRUE,
                        FALSE)) %>%
  select(-flag1)
head(mean_hourly_wage, n = 10)

```
```{r wages-demog-hs}
# join the wages information and the demographic information by case id.
wages_demog <- left_join(mean_hourly_wage, full_demographics, by="id")
# calculate the years in work force and the age of the subjects in 1979
wages_demog <- wages_demog %>%
  mutate(yr_hgc = as.numeric(yr_hgc)) %>%
  mutate(years_in_workforce = year - yr_hgc) %>%
  mutate(age_1979 = 1979 - (dob_year + 1900))
# filter only the id with high school education
wages_before <- wages_demog  %>% filter(grepl("GRADE", hgc))
# calculate the number of observation
keep_me <- wages_before %>%
  count(id) %>%
  filter(n > 4)
wages_before <- wages_before %>%
  filter(id %in% keep_me$id)
```