### 2021-09-28 data-raw/data-preprocessing.Rmd
remove docs from gitignore, add docs

The data cleaning step is conducted using the suite of packages in `tidyverse`.

```{r load-pkgs}
library(tidyverse)
select <- dplyr::select
```

# Reading the data {#read-data}

```{r raw-data, eval=FALSE, include=FALSE}
source(here::here("data-raw/NLSY79/NLSY79.R"))
```

<details class="source">
<summary>Click here for the **source file to read data** (note this is quite long)</summary>
The code below is provided by the NLSY79 database to do the
reading and initial cleaning of data.

```{r raw-data-rmd, code=readLines(here::here("data-raw/NLSY79/NLSY79.R"))}
```

</details>

The above source code creates a data set `new_data_qnames`.
As shown below, the column names contain information on the job number
(HRP1 = job 1, HRP2 = job 2, ..., HRP5 = job 5) and the year information.

```{r untidy-data, echo=-1}
options(width=70)
new_data_qnames %>%
  select(CASEID_1979,
         starts_with("HRP") &
           ends_with(c("1979", "1980")),
         everything()) %>%
  str(list.len = 10)
```

# Demographic data pre-processing
## Tidy the date of birth data
The month and year of birth is recorded in 1979 and 1981 for each individual.
The records in 1981 are missing for some individuals so we take the month and year of
birth from 1979 records.

Where the record is present for both 1979 and 1981, we check that the
record matches.

```{r dob-tidy}
dob_tidy <- new_data_qnames %>%
  select(id = CASEID_1979,
         year_1979 = `Q1-3_A~Y_1979`,
         year_1981 = `Q1-3_A~Y_1981`,
         month_1979 = `Q1-3_A~M_1979`,
         month_1981 = `Q1-3_A~M_1981`) %>%
  mutate(dob_year = case_when(
                      # if the years recorded in both sets match, take 79 data
                      year_1979 == year_1981 ~ year_1979,
                      # if the year in the 81 set is missing, take the 79 data
                      is.na(year_1981) ~ year_1979,
                      # if the sets don't match for dob year, take the 79 data
                      year_1979 != year_1981 ~ year_1979),
         dob_month = case_when(
                      # if months recorded in both sets match, take 79 data
                      month_1979 == month_1981 ~ month_1979,
                      # if month in 81 set is missing, take the 79 data
                      is.na(month_1981) ~ month_1979,
                      # if sets don't match for dob month, take the 79 data
                      month_1979 != month_1981 ~ month_1979),
         # flag if there is a conflict between dob recorded in 79 and 81
         dob_conflict = case_when(
                      !is.na(month_1981) && (month_1979 != month_1981) ~ TRUE,
                      !is.na(year_1981) && (year_1979 != year_1981) ~ TRUE,
                      (year_1979 == year_1981) & (month_1979 == month_1981) ~ FALSE,
                      is.na(month_1981) | is.na(year_1981) ~ FALSE)) %>%
  select(id,
         dob_month,
         dob_year,
         dob_conflict)

has_dob_conflict <- any(dob_tidy$dob_conflict, na.rm = TRUE)
```
```{r dob-conflict-no, echo=FALSE, eval=!has_dob_conflict}
cat("All birth month and year recorded in 1979 and 1981 match.")
```
```{r dob-conflict-yes, echo=has_dob_conflict, eval=has_dob_conflict}
cat("The birth record does not match for the following individuals.")
dob_tidy %>%
  filter(dob_conflict)
```

## Tidy the sex and race data

```{r demog-tidy}
demog_tidy <- categories_qnames %>%
  select(id = CASEID_1979,
         race = SAMPLE_RACE_78SCRN,
         gender = SAMPLE_SEX_1979)
```

## Tidy the grade completed in each year

```{r demog-ed}
demog_education <- new_data_qnames %>%
  # in 2018, the variable's name is Q3-4_2018, instead of HGC_2018
  rename(HGC_2018 = `Q3-4_2018`) %>%
  select(id = CASEID_1979,
         starts_with("HGCREV"),
         "HGC_2012",
         "HGC_2014",
         "HGC_2016",
         "HGC_2018") %>%
  pivot_longer(!id,
               names_to = "var",
               values_to = "grade") %>%
  separate("var", c("var", "year"), sep = "_") %>%
  filter(!is.na(grade)) %>%
  select(-var)
```

### Get the highest year completed

```{r tidy-hgc}
highest_year <- demog_education %>%
  group_by(id) %>%
  mutate(hgc_i = max(grade)) %>%
  filter(hgc_i == grade) %>%
  filter(year == first(year)) %>%
  rename(yr_hgc = year) %>%
  select(id, yr_hgc, hgc_i) %>%
  ungroup() %>%
  mutate(hgc = case_when(hgc_i == 0 ~ "NONE",
                         hgc_i == 1 ~ "1ST GRADE",
                         hgc_i == 2 ~ "2ND GRADE",
                         hgc_i == 3 ~ "3RD GRADE",
                         hgc_i >= 4 & hgc_i <= 12 ~ paste0(hgc_i, "TH GRADE"),
                         hgc_i == 13 ~ "1ST YEAR COL",
                         hgc_i == 14 ~ "2ND YEAR COL",
                         hgc_i == 15 ~ "3RD YEAR COL",
                         hgc_i == 95 ~ "UNGRADED",
                         TRUE ~ paste0((hgc_i - 12), "TH YEAR COL")))

```
```{r full-demog}
full_demographics <- full_join(dob_tidy, demog_tidy, by = "id") %>%
  full_join(highest_year, by = "id") %>%
  mutate(age_1979 = 1979 - (dob_year + 1900))

```
```{r tidy-hours}
# make a list for years where we used the "QES-52A"
year_A <- c(1979:1987, 1993)
# function to get the hour of work
get_hour <- function(year) {
  if(year %in% year_A){
    temp <- new_data_qnames %>%
      select(CASEID_1979,
             starts_with("QES-52A") &
             ends_with(as.character(year)))}
  else{
    temp <- new_data_qnames %>%
      select(CASEID_1979,
             starts_with("QES-52D") &
             ends_with(as.character(year)))}
  temp %>%
    pivot_longer(!CASEID_1979,
                 names_to = "job",
                 values_to = "hours_work") %>%
    separate("job", c("job", "year"), sep = -4) %>%
    mutate(job = paste0("job_", substr(job, 9, 10))) %>%
    rename(id = CASEID_1979)
}

# getting the hours of work of all observations
years <- c(1979:1994, seq(1996, 2018, by = 2))
hours_all <- map_dfr(years, get_hour)

```
```{r tidy-rate}
get_rate <- function(year) {
  new_data_qnames %>%
    select(CASEID_1979,
           starts_with("HRP") &
              ends_with(as.character(year))) %>%
    pivot_longer(!CASEID_1979, names_to = "job", values_to = "rate_per_hour") %>%
    separate("job", c("job", "year"), sep = -4) %>%
    mutate(job = paste0("job_0", substr(job, 4, 4))) %>%
    rename(id = CASEID_1979)
}

rates_all <- map_dfr(years, get_rate)

# join hours and rates variable
hours_wages <- left_join(rates_all,
                         hours_all,
                         by = c("id", "year", "job")) %>%
  # set the 0 value in rate_per_hour as NA
  mutate(rate_per_hour = ifelse(rate_per_hour == 0, NA, rate_per_hour),
         hours_work = ifelse(hours_work == 0 | hours_work > 84, NA, hours_work))

```
```{r tidy-nojob}
# calculate number of jobs that a person has in one year
no_job <- hours_wages %>%
  filter(!is.na(rate_per_hour)) %>%
  group_by(id, year) %>%
  summarise(no_jobs = length(rate_per_hour))

# filter the observations with available rate per hour
eligible_wages <- hours_wages %>%
  filter(!is.na(rate_per_hour)) %>%
  left_join(no_job, by = c("id", "year"))

# calculate the mean_hourly_wage
# flag1 = code 1 for weighted mean
# code 0 for arithmetic mean
mean_hourly_wage <- eligible_wages %>%
  group_by(id, year) %>%
  #calculate the weighted mean if the number of jobs > 1
  mutate(wages = ifelse(no_jobs == 1, rate_per_hour/100,
                        weighted.mean(rate_per_hour, hours_work, na.rm = TRUE)/100)) %>%
  #give the flag if it the weighted mean
  mutate(flag1 = ifelse(!is.na(wages) & no_jobs != 1, 1,
                        0)) %>%
  #calculate the arithmetic mean for the na
  mutate(wages = ifelse(is.na(wages), mean(rate_per_hour)/100,
                        wages)) %>%
  group_by(id, year) %>%
  summarise(wages = mean(wages),
            total_hours = sum(hours_work),
            number_of_jobs = mean(no_jobs),
            flag1 = mean(flag1)) %>%
  mutate(year = as.numeric(year)) %>%
  ungroup() %>%
  rename(mean_hourly_wage = wages) %>%
  mutate(is_wm = ifelse(flag1 == 1, TRUE,
                        FALSE)) %>%
  select(-flag1)
head(mean_hourly_wage, n = 10)

```
```{r wages-demog-hs}
# join the wages information and the demographic information by case id.
wages_demog <- left_join(mean_hourly_wage, full_demographics, by="id")
# calculate the years in work force and the age of the subjects in 1979
wages_demog <- wages_demog %>%
  mutate(yr_hgc = as.numeric(yr_hgc)) %>%
  mutate(years_in_workforce = year - yr_hgc) %>%
  mutate(age_1979 = 1979 - (dob_year + 1900))
# filter only the id with high school education
wages_before <- wages_demog  %>% filter(grepl("GRADE", hgc))
# calculate the number of observation
keep_me <- wages_before %>%
  count(id) %>%
  filter(n > 4)
wages_before <- wages_before %>%
  filter(id %in% keep_me$id)
```