### 2021-09-29 app/app.Rmd

---
title: "Customize Your Threshold"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(plotly)
library(shiny)
library(shinyWidgets)
```

```{r data}
wages_hs <- readRDS(here::here("wages_hs.rds")) %>%
    dplyr::select(-`X1.x`, -`X1.y`)
```

Inputs {.sidebar}
-----------------------------------------------------------------------
<br/>

<style>
div.blue { background-color:#FADBD8; border-radius: 10px; padding: 10px;}
</style>
<div class = "blue">

```{r}
p(icon("info-circle", lib = "font-awesome"), strong("About"))
```

The wages data in `yowie` package uses the weight of robust linear model to detect anomalies and imputes those anomalies with its fitted values. The threshold of weight used is 0.12 to maintain the natural variability of the data. Here, you can customize your own threshold.

</div>

<br/>

```{r}
sliderInput("th", "Slide to set the threshold",
            min = 0.01,
            max = 1,
            value = 5)
```

<br/>

<style>
div.blue { background-color:#FADBD8; border-radius: 10px; padding: 10px;}
</style>
<div class = "blue">

```{r}
p(icon("info-circle", lib = "font-awesome"), strong("More"))
```

Have any issue or question? 
Please put it [here](https://github.com/numbats/yowie/issues)

</div>

Row {data-height=300}
-------------------------------------

### Spaghetti plots

```{r}

thres_data_all <- reactive({
  temp_data <- wages_hs %>%
    mutate(wages_rlm = ifelse(w < input$th  & .fitted >= 0, .fitted,
                                      mean_hourly_wage)) %>%
    mutate(is_pred = ifelse(w < input$th & .fitted >= 0, TRUE, FALSE))

  wages_compare <- temp_data %>%
    select(id, year, mean_hourly_wage, wages_rlm) %>%
    rename(original = mean_hourly_wage,
           imputed = wages_rlm) %>%
    pivot_longer(c(-id, -year), names_to = "type", values_to = "wages") %>%
    mutate(type = factor(type, levels = c("original", "imputed")))
  })

renderPlot({
  ggplot(thres_data_all()) +
    geom_line(aes(x = year,
                  y = wages,
                  group = id),
              alpha = 0.1) +
    theme_bw() +
    facet_wrap(~type, scales = "free_y") +
    theme(strip.text = element_text(size = 12))
  })
```

### Summary Statistics

```{r}
p(icon("info-circle", lib = "font-awesome"), "The imputed plot shows how the data changes after the threshold is customized. You can then compare it with the original plot. The following is the summary statistics of the imputed data.")
```

```{r}
max_ori <- max(wages_hs$mean_hourly_wage)
min_ori <- round(min(wages_hs$mean_hourly_wage), 2)
med_ori <- round(median(wages_hs$mean_hourly_wage), 2)
mean_ori <- round(mean(wages_hs$mean_hourly_wage), 2)
```

```{r}
pred <- reactive({
  temp_data <- wages_hs %>%
    mutate(wages_rlm = ifelse(w < input$th  & .fitted >= 0, .fitted,
                                      mean_hourly_wage)) %>%
    mutate(is_pred = ifelse(w < input$th & .fitted >= 0, TRUE, FALSE))
  })
```

<style>
div.blue { background-color:#FADBD8; border-radius: 10px; padding: 10px;}
</style>
<div class = "blue">

```{r}
renderText({paste0("Minimum = ", round(min(pred()$wages_rlm), 2), " (Original = 0.01)")})
renderText({paste0("Median = ", round(median(pred()$wages_rlm), 2), " (Original = 7.2)")})
renderText({paste0("Mean = ", round(mean(pred()$wages_rlm), 2), " (Original = 11.87)")})
renderText({paste0("Maximum = ", round(max(pred()$wages_rlm), 2), " (Original = 33400)")})
```

</div>

### Box

```{r}
renderValueBox({
  n_pred <- round(nrow(filter(pred(), is_pred == TRUE))/nrow(pred())*100,2)
  valueBox(
    value = n_pred,
    caption = "percentage of imputed observations",
    icon = "fa-area-chart"
  )
})

```

Row {data-height=650}
-------------------------------------

### Individual plots

```{r}
actionButton("resample", "Click here to sample individuals to display")
```

```{r}

compare_data <- eventReactive(input$resample, {
  sample_id <- sample(unique(wages_hs$id), 16)
  sample <- subset(wages_hs, id %in% sample_id)
})

thres_data <- reactive({
  temp_data <- compare_data() %>%
    mutate(wages_rlm = ifelse(w < input$th  & .fitted >= 0, .fitted,
                                      mean_hourly_wage)) %>%
    mutate(is_pred = ifelse(w < input$th & .fitted >= 0, TRUE, FALSE))

  wages_compare <- temp_data %>%
    select(id, year, mean_hourly_wage, wages_rlm) %>%
    rename(original = mean_hourly_wage,
           imputed = wages_rlm) %>%
    pivot_longer(c(-id, -year), names_to = "type", values_to = "wages") %>%
    mutate(type = factor(type, levels = c("original", "imputed")))
  })

renderPlot({
  ggplot(thres_data()) +
    geom_line(aes(x = year,
                  y = wages,
                  colour = type,
                  linetype = type),
              alpha = 1,
              size = 1) +
    geom_point(aes(x = year,
                   y = wages,
                   colour = type),
               alpha = 0.5,
               size = 1) +
    scale_color_manual(values = c("#229954", "#DC7633")) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 10, size = 5),
          legend.position = "bottom",
          legend.text = element_text(size = 12),
          legend.key.size = unit(6,"line")) +
    facet_wrap(~id, scales = "free_y")
  })

```